Distributed Behavior in a Fluid Analogy Architecture

### 2017-12-08 papers/draft.tex
Adds initial probability-adjustment-formula changes

\begin{enumerate}
    \item $H_i$ Centralized variables constrict copycat's ability.
    \item $H_0$ Centralized variables either improve or have no effect on copycat's ability.
\end{enumerate}

@@

    The aim of this paper is to create and test a new version of the copycat software that makes effective use of a multiple level description.
    Until now, copycat has made many of its decisions based on a global variable, \emph{temperature}.
    [].

@@

        Once the effect of temperature was evaluated, new temperature-based probability adjustment formulas were proposed that each had a significant effect on the answer distributions produced by copycat.
        Instead of representing a temperature-less, decentralized version of copycat, these formulas are meant to represent the centralized branch of copycat.
        [Insert formula write-up]

>>>>

\begin{enumerate}
    \item $H_i$ Centralized, global variables constrict copycat's ability.
    \item $H_0$ Centralized, global variables either improve or have no effect on copycat's ability.
\end{enumerate}

@@

    The aim of this paper is to create and test a new version of the copycat software that makes effective use of a multiple level description.
    Until now, copycat has made many of its decisions, even local ones, based on a global variable, \emph{temperature}.
    This paper will evaluate alternatives to this global decision, and compare several different variant of the copycat software to effectively decide on which design choices to make.

@@

        Once the effect of temperature was evaluated, new temperature-based probability adjustment formulas were proposed that each had a significant effect on the answer distributions produced by copycat.
        Instead of representing a temperature-less, decentralized version of copycat, these formulas are meant to represent the centralized branch of copycat.

        These formulas curve probabilities, making unlikely events more likely and likely events less likely as a function of the global \emph{temperature} variable.

        The desired (LISP documented) behavior is as follows:
        At high temperatures, the system should explore options that would otherwise be unlikely.
        So, at temperatures above half of the maximum temperature, probabilities with a base value less than fifty percent will be curved higher, to some threshold.
        At temperatures below half of the maximum temperature, probabilities with a base value above fifty percent will be curved lower, to some threshold.

        The original formulas being used to do this were overly complicated.
        In summary, many formulas were tested in a spreadsheet, and an optimal one was chosen that replicated the desired behavior.
        The remainder of the section discusses different formulas and their advantages/disadvantages.
        Also, as a general rule, changing these formulas causes copycat to produce statistically significantly different answer distributions.

        The original formula for curving probabilties in copycat:
        \lstinputlisting[language=Python]{formulas/original.py}

        An alternative that seems to improve performance on the "abd:abd::xyz:\_" problem:
        This formula produces probabilities that are not bounded between 0 and 1. These are generally truncated.
        \lstinputlisting[language=Python]{formulas/entropy.py}

        However, this formula worsens performance on non "xyz" problems.
        Likely, because of how novel the "xyz" problem is, it will require more advanced architecture changes.
        For instance, MetaCat claims to assist in solving the "xyz" problem.

        The entropy formula is an improvement, but other formulas are possible too.

        Below are variations on a "weighted" formula.
        The general structure is:

        \[\emph{p'} = \frac{T}{100} * S + \frac{100-T}{100} * U\]

        Where: $S$ is the convergence value for when $T = 100$ and
               $U$ is the convergence value for when $T = 0$.
        The below formulas simply experiment with different values for $S$ and $U$

        \lstinputlisting[language=Python]{formulas/weighted.py}

        After some experimentation and reading the original copycat documentation, it was clear that $S$ should be chosen to be $0.5$ (All events are equally likely at high temperature) and that $U$ should implement the probability curving desired at low temperatures.

        The following formulas let $U = p^r$ if $p < 0.5$ and let $U = p^\frac{1}{r}$ if $p >= 0.5$.
        This controls whether/when curving happens.
        Now, the \emph{single} parameter $r$ simply controls the degree to which curving happens.
        Different values of $r$ were experimented with (values between $10$ and $1$ were experimented with at increasingly smaller step sizes).
        $2$ and $1.05$ are both good choices at  opposite "extremes".
        $2$ works because it is large enough to produce novel changes in behavior at extreme temperatures without totally disregarding the original probabilities.
        Values above $2$ do not work because they make probabilities too uniform.
        Values below $2$ (and above $1.05$) are feasible, but produce less curving and therefore less unique behavior.
        $1.05$ works because it very closely replicates the original copycat formulas, providing a very smooth curving.
        Values beneath $1.05$ essentially leave probabilities unaffected, producing no significant unique behavior dependent on temperature.

        \lstinputlisting[language=Python]{formulas/best.py}

        All of these separate formulas will later be cross-compared to other variants of the copycat software using a Pearson's $\chi^2$ test.
